---
title: "R Notebook"
output: html_notebook
---
# getting a basic summary for the project dataset






```{r}
library(ggplot2)
library(dplyr)
library(statsr)
library(gridExtra)
library(knitr)
```


```{r}
install.packages("statsr")
```

```{r}
Life_Expantancy_data <- read.csv("C:/Users/yuqia/Desktop/Life Expectancy Data.csv")
```
```{r}
Life_Expantancy_data
```

```{r}
head(Life_Expantancy_data)
```
#getting the basic summary of the datset
```{r}
summary(Life_Expantancy_data)
```

```{r}
# remove na in r - remove rows - na.omit function / option
clean_data <- na.omit(Life_Expantancy_data) 
```


```{r}
clean_data
```




#Exploratory data Analysis
#1. checking the distribution of target variable
```{r}
library(ggplot2)

```
#Building the histogram 
```{r}
ggplot(data=clean_data,aes(clean_data$Life.expectancy))+geom_histogram(aes(y=..density..),fill="orange")+geom_density()
```

#Checking distribution for the predictor variables

```{r}
ggplot(data=clean_data,aes(clean_data$Alcohol))+geom_histogram(aes(y=..density..),fill="orange")+geom_density()
```
#getting the distribuition for adult morality

```{r}
ggplot(data=clean_data,aes(clean_data$Adult.Mortality))+geom_histogram(aes(y=..density..),fill="orange")+geom_density()
```
#getting the distribution for infant death

```{r}
ggplot(data=clean_data,aes(clean_data$infant.deaths))+geom_histogram(aes(y=..density..),fill="orange")+geom_density()
```
#getting distribution for HIV
```{r}
ggplot(data=clean_data,aes(clean_data$HIV.AIDS))+geom_histogram(aes(y=..density..),fill="orange")+geom_density()
```
#getting distribution for Polio
```{r}
ggplot(data=clean_data,aes(clean_data$Polio))+geom_histogram(aes(y=..density..),fill="orange")+geom_density()
```
#getting distibution for thinness..10 to 19.years

```{r}
ggplot(data=clean_data,aes(clean_data$thinness..1.19.years))+geom_histogram(aes(y=..density..),fill="orange")+geom_density()
```
#getting distribution for percentage expenditure
```{r}
ggplot(data=clean_data,aes(clean_data$percentage.expenditure))+geom_histogram(aes(y=..density..),fill="orange")+geom_density()
```
#getting distribution for under five death

```{r}
ggplot(data=clean_data,aes(clean_data$under.five.deaths))+geom_histogram(aes(y=..density..),fill="orange")+geom_density()
```

#According to the above graph, the Life_expantancy variable follow a little bit sked to the left distributions.it is not as normal as I expepected.

# check correlations of the target variable with the first 5 predictors using Pearson correlation
```{r}
library(psych)
```

```{r}

pairs.panels(clean_data[,4:9], 
             method = "pearson", # correlation method
             hist.col = "green",
             density = TRUE,  # show density plots
             ellipses = TRUE # show correlation ellipses
)

```
```{r}
pairs.panels(clean_data[,c(4,10:14)], 
             method = "pearson", # correlation method
             hist.col = "green",
             density = TRUE,  # show density plots
             ellipses = TRUE # show correlation ellipses
)
```
```{r}
pairs.panels(clean_data[,c(4,15:19)], 
             method = "pearson", # correlation method
             hist.col = "green",
             density = TRUE,  # show density plots
             ellipses = TRUE # show correlation ellipses
)

```
```{r}
pairs.panels(clean_data[,c(4,20:22)], 
             method = "pearson", # correlation method
             hist.col = "green",
             density = TRUE,  # show density plots
             ellipses = TRUE # show correlation ellipses
)
```
#following the rule
#Negligible correlation = 0.00 - 0.09
#Weak correlation = 0.10 - 0.39
#Moderate correlation = 0.40 - 0.69
#Strong correlation = 0.70 - 0.89
#Very strong correlation = 0.90 - 1.00



# By checking the correlation, we can observe that income.composition of resources, scholling, and adult morality have a strong correlation iwth the target variable life exptancy

#By checking the correlation, we can observe that BMI, HIV.AIDS, Diphtheria, thinness..1.19.years, Polio, thinness.5.9.years, and GDP have a moderate correlation with the target variable life exptancy

#By checking the correlation, we can observe that Alcohol, percentage.expenditure, under.five.deaths, Total.expenditure, and infant.deaths have a weak correlation with the target variable life exptancy,

#By checking the correlation, we can observe that Hepatitis.B, Measles, and Population have extremely weak correlation with the target variable life exptancy.




#Doing modleing and predictions
# First steps: We should oberve the null model
#building the null model
```{r}
nullmodel<-lm(Life.expectancy~1,data=clean_data)
```

#Getting the summary of the null model
```{r}
summary(nullmodel)
```


#make predictions
```{r}
nullmodelpredicts<-predict(nullmodel,  # my model
                    newdata = clean_data, # dataset
                    type = "response") # to get predicted valu
```


```{r}
nullmodelpredicts
```
#Second, let us observe the full model
#Building the full model
```{r}
fullmodel<-lm(Life.expectancy~.,data=clean_data)
```
#getting the summary of the full model
```{r}
summary(fullmodel)
```

#Make prediction based on the full model
```{r}
fullmodelPreds <- predict(fullmodel,  # my model
                    newdata = clean_data, # dataset
                    type = "response")

```
#Getting the predictions
```{r}
fullmodelPreds
```
#doing the backward step model
```{r}
BackwardStepModel <- step(fullmodel,
                      direction = "backward")
```
#getting the summary of the backwardstepmodel
```{r}
summary(BackwardStepModel)
```





#Make the predictions for the backwardstep 
```{r}
BackwardStepModelPreds <- predict(BackwardStepModel,  # my model
                    newdata = clean_data, # dataset
                    type = "response") # to get predicted values
```

#show the backward predictions
```{r}
BackwardStepModelPreds
```
#Doing the forward model selections

```{r}
forwardStepModel <- step(nullmodel,
                         direction="forward",
                         scope=list(lower=nullmodel,
                                    upper=fullmodel))
```

#Getting the summary of the forward method
```{r}
summary(forwardStepModel)
```
#make the predictions
```{r}
forwardStepModelPreds <- predict(forwardStepModel,  # my model
                    newdata = clean_data, # dataset
                    type = "response") # to get predicted values
```

#showing the prediction
```{r}
forwardStepModelPreds
```

#Checking the assumption for normality fr different models
```{r}
par(mfrow = c(3, 2))
plot(density(fullmodel$residuals))
plot(density(nullmodel$residuals))
plot(density(BackwardStepModel$residuals))
plot(density(forwardStepModel$residuals))

```
#checking the null model assumption plot
```{r}
par(mfrow = c(2, 2))
plot(nullmodel)
```
#checking assumtion for the full model
```{r}
par(mfrow = c(2, 2))
plot(fullmodel)
```
#checking the assumption for the
```{r}
par(mfrow = c(2, 2))
plot(BackwardStepModel)
```


#checking assumotion for the forward model
```{r}
par(mfrow = c(2, 2))
plot(forwardStepModel)
```
# Getting the adjusted R-squared for different model
```{r}
summary(fullmodel)$adj.r.squared
summary(nullmodel)$adj.r.squared 
summary(BackwardStepModel)$adj.r.squared 
summary(forwardStepModel)$adj.r.squared 

```
#loading the library
```{r}
library(Matrix)
```


# The backward methods is always the key, decided to make transformation to normalized that model
```{r}
original_model<-Life.expectancy ~ Country + Year + infant.deaths + Alcohol + 
    Hepatitis.B + under.five.deaths + HIV.AIDS + thinness.5.9.years + 
    Income.composition.of.resources + Schooling
```


```{r}
new_model1<-lm(log(Life.expectancy )~ Country + Year + infant.deaths + log(Alcohol) + 
    Hepatitis.B + under.five.deaths + HIV.AIDS + thinness.5.9.years + 
    Income.composition.of.resources + Schooling,data=Life_Expantancy_data)
```

```{r}
new_model1
```

```{r}
summary(new_model1)
```

```{r}
par(mfrow = c(2, 2))
plot(new_model1)
```


#First, Identify outliers
```{r}
boxplot(clean_data$Life.expectancy,plot=FALSE)$out
```
#Save the outliers in a vectors
```{r}
outliers <- boxplot(clean_data$Life.expectancy, plot=FALSE)$out

```

```{r}
install.packages("CRAN")

```

```{r}
library(CRAN)
```

```{r}
clean_data[which(clean_data$Life.expectancy %in% outliers),]
```
# removing the row contain outliers
```{r}
clean_data1 <- clean_data[-which(clean_data$Life.expectancy %in% outliers),]
```

```{r}
boxplot(clean_data$Life.expectancy)
```



```{r}
new_model2<-lm(Life.expectancy~ Country + Year + infant.deaths + Alcohol + 
    Hepatitis.B + under.five.deaths + HIV.AIDS + thinness.5.9.years + 
    Income.composition.of.resources + Schooling,data=clean_data1)
```


```{r}
par(mfrow = c(2, 2))
plot(new_model2)
```
# USing only the model who have strong correlations with the response variable life.Exptancy
```{r}

model3 <- lm(Life.expectancy ~ Schooling +  Adult.Mortality + Income.composition.of.resources,
                data = clean_data)


```

#getting the summary for model3
```{r}
summary(model3)
```


```{r}
par(mfrow = c(2, 2))
plot(model3)
```


```{r}
Model3Preds <- predict(model3,  # my model
                    newdata = clean_data, # dataset
                    type = "response") # to get predicted values
```

#showing the prediction of model 3
```{r}
Model3Preds
```
# lastly, let us check the colinearity fr different model

#check collinearity of the backward Model
```{r}
library(car)
```

```{r}
vif(BackwardStepModel)
```

#checking the collineairty for the forward model
```{r}
vif(forwardStepModel)
```
#Checking the collineairty for the model3
```{r}
vif(model3)
```

#Getting the Anova table for model 3
```{r}
anova(lm(Life.expectancy ~ Schooling + Adult.Mortality + 
    Income.composition.of.resources, data = clean_data))

```
```{r}
install.packages("sjPlot")
```

```{r}
install.packages("sjmisc")
```

```{r}
install.packages("sjlabelled")
```

```{r}
library(sjPlot)
library(sjmisc)
library(sjlabelled)

```



```{r}
tablemodel<-tab_model(model3)
```


```{r}
tablemodel
```












































































```

